---
title: "Tic-Tac-Toe"
author: "Johan Jordaan"
date: "14/12/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("dplyr")
library("ggplot2")
library("RColorBrewer")
```

## Introduction
I love playing all sorts of games and it seemed like a natural place to apply some of my machine learning and data science knowledge. I decided to start with a all time classic, Tic-Tac-Toe. I reasoned that this will allow me to concentrate on developing a methodology for investigating the properties of games without concentradting to much on the actual game.  

## Methodology
I am assuming that I don't know the actual game or it's rules.Obviously I know that the game is Tic-Tac-Toe, but as far as possible I will edevour to not let my analysis be clouded by what I know of TTT. All I have access to is a set of simulations of the game. Each line item contains the following data:

Game specific elements
* gameId
* numMoves
* winner

Move specific elements
* moveNumber
* result
* move
* player

## New players
I am going to develop some new players based on the anayslsis.

# Two random players

## Initial analysis
```{r load_data,echo=FALSE}
d <- readRDS("./moves100k.rds")
numMovesStats = d %>% filter(moveNumber==1) %>% summarise(median=median(numMoves), mean=mean(numMoves),sd=sd(numMoves)) 
```

Total number of games ```r length(unique(d$gameId))```

Total number of moves ```r nrow(d)```

Number of players ```r length(unique(d$player))``` [```r levels(d$player)```]

Number of game state ```r length(unique(d$winner))``` [```r levels(d$winner)```]

## Basic stats

Average number of moves per game ```r numMovesStats$mean``` with a sd of ```r numMovesStats$sd```

Minimum number of moves per game ```r min(d$numMoves)```

Starting players seems to be randoly distributed:
```{r staring_player_stats,echo=FALSE}
d_startingPlayer <- d %>% group_by(startingPlayer) %>% summarise(cnt=n())
print(d_startingPlayer)
bt1 <- binom.test(d_startingPlayer[1,]$cnt,d_startingPlayer[1,]$cnt+d_startingPlayer[2,]$cnt)
print(bt1)
```

Looking at the number of wins per user it seem like wins and losses are randomly distributed between the two players.
```{r game_stats,echo=FALSE}
stats = d %>% 
        filter(moveNumber==1) %>% 
        group_by(winner) %>% 
        summarise(count=n())
print(stats)
```


A binomial test confirms that the probability of winning and losing a random game is 50%.
```{r win_loss_random,echo=FALSE}
wins <- stats$count[1]
losses <- stats$count[2]
print(binom.test(wins,wins+losses))

gd <- d %>% 
     filter(moveNumber==1) %>% 
     group_by(winner) 

g0 <- ggplot(data=gd,aes(x=winner,fill=winner))+ geom_bar()+scale_fill_brewer(palette = "Accent")
print(g0)
```

## Move Number Analysis
I found the following graph very interresting:
```{r move_wl_graph,echo=FALSE}
g1 <- ggplot(data=d,aes(x=moveNumber,fill=result))+
      geom_bar()+
      scale_x_discrete(name ="Move Number", limits=unique(d$moveNumber) )+
      scale_fill_brewer(palette = "Accent") 
print(g1)
```

It's a plot of the number of eventual wins, losses and draws resulting from a particular tien number. Where turn number here is a sinomym for move number. From this graph these interresting observations pop up:

1. It seems like the player who playes first has a distinct advantage.  
2. After turn 5 the first player's advantage increases 
3. After turn 5 the probability of a draw increases.

Lets look at this data but in terms of probabilities of winning and losing:

```{r move_wl_last4_turns,echo=FALSE}
d2 <- d %>% group_by(moveNumber) %>% summarize(count=n()) %>% inner_join(d,by="moveNumber") %>% group_by(moveNumber,result) %>% summarise(perc=n()/unique(count))

g2 <- ggplot(data=d2,aes(x=moveNumber,y=perc,color=result))+
      geom_point(alpha=0.5) +
      scale_x_discrete(name ="Move Number", limits=unique(d$moveNumber) )+
      scale_color_brewer(palette = "Accent")  +
      geom_smooth(method="loess",se=FALSE)
print(g2)
```

This graph shows the obesrvation much more clearly.

## Move analysis
Now lets look at moves idivisuallin outside of te context of a particular game. The firts thing that comes to mind is a graph of how many times a particular move led to a particular outcome:

```{r move_outcome_graph,echo=FALSE}
d3 <- d %>% group_by(move) %>% 
      summarize(count=n()) %>% 
      inner_join(d,by="move") %>% 
      group_by(move,result) %>% 
      summarise(perc=n()/unique(count))
g3 <- ggplot(data=d3,aes(x=move,y=perc,fill=result))+
      geom_bar(stat="identity")+
      scale_x_discrete(name ="Move", limits=c(1:9) )+
      scale_fill_brewer(palette = "Accent") 
print(g3)
```

What a beautiful graph. It shows us exactly which moves has the highest probabilty of leading to a win. It seems like move 5 should be the first choice follwed by 1,3,7 and 9 in no particular order and then 2,4,6 and 8 in no particular order.

## Recap
We know now that the staring player has an advantage. We also know that selecting move 5 provides an advantage. My suspicion is that the staring player has an advantage because they will always *assumes prior knowledge*  be able to select move 5. 








## Todo
* Fix the graph lables - Specifically Y axis and legend 
* Maybe remove player X and O to remove prior knowledge bais

## Questions

* Given two random players
* Do you have an advantage if you start
  Null Hypothesis : Both players have same probabilty of winning

* What is your probablity of winning if you start 



